---
title: "Sector growth"
author: "Time series forecasting" 
output:
  html_document:
    css: custom-release.css
    df_print: paged
    number_sections: no
    toc: yes
    toc_float: yes
  html_notebook:
    css: custom-release.css
    toc: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
home_dir <- "~/R scripts/"  # "~/Downloads/Dropbox/Werk/R\ Scripts/" "~/R scripts/"   
setwd(paste0(home_dir, "/R_time_series/"))
library(ggraph) 
library(igraph)

source("project.R")
source("load_companies.R")
source("roll_up_nace_tree.R")
source("forecast_functions.R")

# Configuration file
library(yaml)
config <- read_yaml("config.yml")

# Set up project
open_project("R_time_series", home_dir)

# Set analysis variables
var_date_from <- as.Date(config$date_start) # Start date of time series
var_date_to <- as.Date(config$date_end)     # End date of time series
months_forecast <- config$months_forecast   # Number of months in forecast
```

# Loading and preparing data

The used configuration file, config.yml:
```{r Show config.yml file, echo = FALSE}
cat(readLines("config.yml"), sep = '\n')
```

The company data and SBI code hierarchy is loaded:
```{r Load data, collapse=TRUE}
# Import & transform company data ----
tbl_companies <- prep_companies(config$process_companies, 
                                var_date_from, 
                                var_date_to)
# Import branche hierarchy data
tbl_nace <- read.csv2(paste0(dir_input, "/branche_hierarchy.csv"), 
                      stringsAsFactors = FALSE)
```

## Rolling up the NACE hierarchy

The number of companies per nace code in the company file can be small, making it unsuitable for time series forecasting. For this reason the number of companies are aggregated though the NACE hierarchy, by propagating codes with the small number of companies to higher hierarchy codes.

To get this mechanism working the first need to add the number of companies for each of the NACE codes to the complete NACE hierarchy.
```{r Add the number of companies to the nace hierarchy}
# Add number of companies to the NACE hierarchy, so all 
tbl_nace_qty <- tbl_nace %>% 
  left_join(tbl_companies, by = c("code" = "sbi_full")) %>% 
  group_by(code, code_parent, layer_no) %>% 
  summarise(qty = n_distinct(id_giant, na.rm = TRUE)) %>% 
  ungroup() 
```

This is how the NACE hierarchy looks now; many sectors with a small number of companies:
```{r Display NACE hierarchy before roll-up, echo = FALSE, warning = FALSE}
tbl_nodes <- data_frame(code = with(tbl_nace_qty, unique(c(code, code_parent)))) %>% 
  left_join(tbl_nace_qty, by = "code") %>% 
  rename(code_sector = code)
tbl_links <- tbl_nace_qty

tbl_nodes %<>% mutate(layer_no = factor(layer_no))

graph <- graph_from_data_frame(tbl_links, tbl_nodes, directed = TRUE)

# Create Graph
set.seed(42)
ggraph(graph, layout = "nicely") +
  geom_edge_link(edge_width = 0.5, alpha = .4) +
  geom_node_point(aes(colour = layer_no, size = qty), 
                  alpha = 0.4) +
  # geom_node_text(aes(label = code_sector, colour = layer_no), 
  #                repel = TRUE, size = 10) +
  guides(col = FALSE, size = FALSE) +
  scale_color_graydon() +
  theme_graydon("blank")

rm(graph, tbl_nodes, tbl_links)
```

Now the actual roll up is done after which the 'translation table' is joined to the table with companies table:
```{r Rolling up NACE hierarchy}
lst_nace_recoding <- roll_up_nace_tree(tbl_nace_qty, config$qty_companies_in_rolledup)

tbl_companies %<>% 
  left_join(lst_nace_recoding$tbl_dictionary, by = c("sbi_full" = "code"))
```

That results in the hierarchy below, containing far less sectors with a higher number of companies, making forecasting success more likely:
```{r Plot of rotbllled up NACE hierarchy, echo = FALSE, warning = FALSE}
tbl_nace_qty <- tbl_nace %>% 
  left_join(lst_nace_recoding$tbl_rolled_up, by = "code") %>% 
  select(code, code_parent, layer_no = layer_no.x, qty = qty_sticky)

tbl_nodes <- data_frame(code = with(tbl_nace_qty, unique(c(code, code_parent)))) %>% 
  left_join(tbl_nace_qty, by = "code") %>% 
  rename(code_sector = code)
tbl_links <- tbl_nace_qty

tbl_nodes %<>% mutate(layer_no = factor(layer_no))

graph <- graph_from_data_frame(tbl_links, tbl_nodes, directed = TRUE)

# Create Graph
set.seed(42)
ggraph(graph, layout = "nicely") +
  geom_edge_link(edge_width = 0.5, alpha = .4) +
  geom_node_point(aes(colour = layer_no, size = qty), 
                  alpha = 0.4) +
  # geom_node_text(aes(label = code_sector, colour = layer_no), 
  #                repel = TRUE, size = 10) +
  guides(col = FALSE, size = FALSE) +
  scale_color_graydon() +
  theme_graydon("blank")
```
```{r Clean up stuff, echo=FALSE}
rm(graph, tbl_nodes, tbl_links)
```

## Spreading companies through time

A time series is a series of data points indexed in time order. In  this step the companies are put in each month within a time frame, specified by the config file, according to their date of establishment and discontinuation.

Depending on the processing configuration, the data is processed *or* pre-processed is loaded; after which NACE descriptions are added.
```{r Aggregate the number of companies and time-spread}
# Depending on the processing configuration the data is processed
if(config$process_aggregate){
  tbl_companies_aggr <- aggregate_companies(tbl_companies, lst_nace_recoding$tbl_dictionary)
  saveRDS(tbl_companies_aggr, file = paste0(dir_input, "/companies_aggr.RDS"))
} else {
  # or pre-processed is loaded
  tbl_companies_aggr <- read_rds(paste0(dir_input, "/companies_aggr.RDS"))
}

tbl_companies_aggr %<>%
  rename(code = code_new) %>% 
  left_join(tbl_nace, by = "code")
```

# Initial exploration

## Development of the number of companies per sector

Now we can take a quick look at the development number of companies in each sector:
```{r Plot number of companies per NACE code by month}
ggplot(tbl_companies_aggr, aes(x = month, y = qty_companies)) +
  geom_area(col = col_graydon[2], fill = col_graydon[2], alpha = .6) +
  facet_wrap(~ code) +
  scale_y_continuous(labels = format_number) +
  labs(x = "", y = "# Companies") +
  theme_graydon("grid")
```
The sector with NACE code `r config$code_nace` looks like an interesting candidate, because it has no obvious trend, which would make it more challenging to forecast. 

## Development of retail sector

When zooming in on this NACE code we see how interesting this is in terms of rises and drops in number of companies:
```{r pressure, echo=FALSE}
tbl_companies_aggr %>% 
  filter(code == config$code_nace) %>% 
ggplot(aes(x = month, y = qty_companies)) +
  geom_area(col = col_graydon[2], fill = col_graydon[2], alpha = .6) +
  facet_wrap(~ description) +
  scale_y_continuous(labels = format_number) +
  labs(x = "", y = "# Companies") +
  theme_graydon("grid")
```

# Creating and cleaning the time series

The library **forecast** contains all kinds of time serie goodies and forecasting functions. The **ggfortify** library provides some stuff for plotting and converting time series.
```{r Loading time series packages}
library(forecast)
library(ggfortify)
```

First, the data of the sector we want to explore and forecast is set apart:
```{r Getting specific nace code}
tbl_companies_code <- tbl_companies_aggr %>% filter(code == config$code_nace)
var_sector <- (tbl_nace %>% filter(code == config$code_nace))$description
```

Then a *ts* object is created from this data, by the *ts()* function. The *ts* object is created by :

* a vector with the quantities per month (in this case), 
* the *frequency* i.e. number of units per year (in this case 12, for months) and 
* the date from which the time series should start

```{r Create time series}
ts_companies <- ts(tbl_companies_code$qty_companies, 
                   frequency = 12, 
                   start = c(year(var_date_from), month(var_date_from)) )
```

The time series is cleaned 
```{r Cleaning time series}
ts_companies_clean <- tsclean(ts_companies)
```

### Compare original and cleaned time series
```{r Compare original and cleaned time series}
plot_time_series(list(ts_companies, ts_companies_clean),
                 c("Read", "Cleaned"))
```

## Forecasting

### Create a training set
```{r Create a training set}
ts_companies_train = subset(ts_companies_clean, end = length(ts_companies_clean) - months_forecast)
```

### Create forecast models
```{r Create forecast models}
fit_mean <- meanf(ts_companies_train, h = months_forecast)    # Mean forecasting (mean of all observations)
fit_naive <- naive(ts_companies_train, h = months_forecast)   # Naive
fit_snaive <- snaive(ts_companies_train, h = months_forecast) # Naive seasonal
fit_ses <- ses(ts_companies_train, h = months_forecast)       # Simple exponential smoothing
fit_holt <- holt(ts_companies_train, h = months_forecast)     # Holt’s linear trend
fit_hw_m <- hw(ts_companies_train, h = months_forecast, seasonal = "multiplicative")  # Holt-Winters multiplicative
fit_hw_a <- hw(ts_companies_train, h = months_forecast, seasonal = "additive")        # Holt-Winters additive
model_ets <- ets(ts_companies_train)                # Errors, Trend, and Seasonality (ETS) 
fit_ets <- forecast(model_ets, h = months_forecast)
model_auto.arima <- auto.arima(ts_companies_train)  # ARIMA (automatic)
fit_auto.arima <- forecast(model_auto.arima, h = months_forecast)
```

### Evaluate forecast models by MASE
```{r Evaluate forecast models}
lst_fitted <- list(fit_mean, fit_naive, fit_snaive, fit_holt, fit_hw_m, fit_hw_a, 
                   fit_ets, fit_ses, fit_auto.arima)
lst_evaluations <- lapply(lst_fitted, evaluate_forecast, ts_companies_clean) 
```

### All forecast model results
```{r Plot all forecast methods, warning=FALSE}
lst_forecast_plots <- sapply(lst_evaluations, "[", 1)
do.call("grid.arrange", c(lst_forecast_plots, ncol=3))
```

### Compare models based on MASE evaluation
```{r Plot MASE of all forecast models}
# Review MASE
tbl_forecast_mase <- do.call("rbind", sapply(lst_evaluations, "[", 2))
tbl_forecast_mase %>% 
  mutate(is_best = MASE == min(MASE)) %>% 
  ggplot(aes(x = reorder(method, MASE), y = MASE)) +
  geom_col(aes(fill = is_best)) +
  geom_text(aes(label = round(MASE, 2)), 
            hjust = -.1) +
  scale_fill_graydon() +
  coord_flip() +
  labs(x = "") +
  guides(fill = FALSE) +
  theme_graydon("vertical")
```

### The best performing forecast method
```{r Plot of the best performing forecast method}
no_best <- (tbl_forecast_mase %>% 
              mutate(row_no = row_number()) %>% 
              mutate(is_best = MASE == min(MASE)) %>% 
              filter(is_best))$row_no 
lst_forecast_plots[no_best] 
```

### Apply best forecasting method
```{r Use best method forecast for longer period}
fit_hw_a <- hw(ts_companies_train, h = 60, seasonal = "additive")
result_fit <- evaluate_forecast(fit_hw_a, ts_companies_clean)
result_fit$p_forecast
```

